---
title: "Ranking Analysis Graph"
author:
date: 
output: pdf_document
---

```{r setup, warning=FALSE, message=FALSE}
## Download Packages
library(readxl)
library(rio)
library(stringr)
library(dplyr)
library(tidyr)
library(countrycode)
library(ggplot2)
library(gridExtra)
library(grid)
library(lattice)
```
Import data from GitHub repository. See [LINK](https://github.com/Global-Development-and-Politics-of-Data/Chapter8_Gender).
```{r}
RA_clean = read.csv("RankingAnalysis_clean.csv") 

# "Syrian Arab Republic" needs to be shortened to accommodate the data visualization
RA_clean$country = gsub("Syrian Arab Republic", "Syria", RA_clean$country) 
```

Create Plot 1: "Top 5 Global Gender Gap Scores by Country (2018)"
```{r}
#PLOT 1 - Top Five GGG Scores
plot1 = (RA_clean %>%
  mutate(GGGRank = factor(ggg_rank18), Country = factor(country)) %>%
  group_by(GGGRank, Country) %>%
  group_by(Country) %>%
  arrange(desc(ggg_2018)) %>%
  head(5)%>%
  ggplot(aes(x = reorder(Country, ggg_2018), y= ggg_2018, fill = GGGRank)) +
  geom_bar(position = 'dodge', stat = 'identity', show.legend = FALSE, fill = "rosybrown2")+
  geom_text(aes(label=ggg_2018),position=position_dodge(width=0.9),
            vjust=-0.25, 
            size = 3, 
            family = "serif")+
  ggtitle("Top 5 Global Gender Gap Scores \nby Country (2018)")+
  theme(plot.title = element_text(hjust = 0.5))+
  xlab("Country")+
  ylab("Global Gender Gap Score (0-1)")+
  theme_minimal()
  )
p1 = plot1 + scale_y_continuous(limit = c(0,1), breaks = seq(0,1,.05)) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 0.20, vjust = .5)) +
  theme(axis.title.x = element_text(colour = "black")) +
  theme(axis.title.y = element_text(colour = "gray10")) +
  theme(plot.title = element_text(hjust = 0.5, colour = "midnightblue", face = "bold"))+
  theme(text=element_text(family="serif")) +
  labs(fill = "Rank #", colour = "gray10")
p1
```

Create Plot 2: "Bottom 5 Global Gender Gap Scores by Country (2018)"
```{r}
#PLOT 2 - Bottom Five GGG Scores
plot2 = (RA_clean %>%
  mutate(GGGRank = factor(ggg_rank18), Country = factor(country)) %>%
  group_by(GGGRank, Country) %>%
  group_by(Country) %>%
  arrange(ggg_2018) %>%
  head(5)%>%
  ggplot(aes(x = reorder(Country, ggg_2018), y= ggg_2018, fill = GGGRank)) +
  geom_bar(position = 'dodge', stat = 'identity', show.legend = FALSE, fill = "rosybrown2")+
  geom_text(aes(label=ggg_2018), position=position_dodge(width=0.9), 
            vjust=-0.25, 
            size = 3, 
            family = "serif")+
  ggtitle("Bottom 5 Global Gender Gap Scores \nby Country (2018)")+
  theme(plot.title = element_text(hjust = 0.5))+
  xlab("Country")+
  ylab("Global Gender Gap Score (0-1)")+
  theme_minimal()
)
p2 = plot2 + scale_y_continuous(limit = c(0,1), breaks = seq(0,1,.05)) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 0.20, vjust = .5)) +
  theme(axis.title.x = element_text(colour = "black")) +
  theme(axis.title.y = element_text(colour = "gray10")) +
  theme(plot.title = element_text(hjust = 0.5, colour = "midnightblue", face = "bold"))+
  theme(text=element_text(family="serif")) +
  labs(fill = "Rank #", colour = "gray10")
p2
```

Create Plot 3: "Top 5 Gender Inequality Index Scores by Country (2018)"
```{r}
# PLOT 3 - Top Five GII Scores
plot3 = (RA_clean %>%
  mutate(GIIRank = factor(gii_rank18), Country = factor(country)) %>%
  group_by(GIIRank, Country) %>%
  group_by(Country) %>%
  arrange(gii_2018)%>%
  head(5)%>%
  ggplot(aes(x = reorder(Country, gii_2018), y= gii_2018, fill = GIIRank))+
  geom_bar(position = 'dodge', stat = 'identity', show.legend = FALSE, fill = "plum")+
  geom_text(aes(label=gii_2018), position=position_dodge(width=0.9), 
            vjust=-0.25, 
            size = 3, 
            family = "serif")+
  ggtitle("Top 5 Global Inequality Index Scores \nby Country (2018)")+
  theme(plot.title = element_text(hjust = 0.5))+
  xlab("Country")+
  ylab("Global Inequality Index Score (0-1)")+
  theme_minimal()
)
p3= plot3 + scale_y_continuous(limit = c(0,1), breaks = seq(0,1,.05)) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 0.20, vjust = .5)) +
  theme(axis.title.x = element_text(colour = "black")) +
  theme(axis.title.y = element_text(colour = "gray10")) +
  theme(plot.title = element_text(hjust = 0.5, colour = "midnightblue", face = "bold"))+
  theme(text=element_text(family="serif")) +
  labs(fill = "Rank #", colour = "gray10")
p3
```

Create Plot 4: "Bottom 5 Global Inequality Index Scores by Country (2018)"
```{r}
#PLOT 4 - Bottom Five GII Scores
plot4 = (RA_clean %>%
  mutate(GIIRank = factor(gii_rank18), Country = factor(country)) %>%
  group_by(GIIRank, Country) %>%
  group_by(Country) %>%
  arrange(desc(gii_2018))%>%
  head(5)%>%
  ggplot(aes(x = reorder(Country, gii_2018), y= gii_2018, fill = GIIRank))+
  geom_bar(position = 'dodge', stat = 'identity', show.legend = FALSE, fill = "plum")+
  geom_text(aes(label=gii_2018), position=position_dodge(width=0.9), 
            vjust=-0.25, 
            size = 3, 
            family = "serif")+
  theme(plot.title = element_text(hjust = 0.5))+
  ggtitle("Bottom 5 Global Inequality Index Scores \nby Country (2018)")+
  xlab("Country")+
  ylab("Global Inequality Index Score (0-1)")+
  theme_minimal()
) 
p4 = plot4 + scale_y_continuous(limit = c(0,1), breaks = seq(0,1,.05)) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 0.20, vjust = .5)) +
  theme(axis.title.x = element_text(colour = "black")) +
  theme(axis.title.y = element_text(colour = "gray10")) +
  theme(plot.title = element_text(hjust = 0.5, colour = "midnightblue", face = "bold"))+
  theme(text=element_text(family="serif")) +
  labs(fill = "Rank #", colour = "gray10")
p4
```

## Combine Plots
Create a combined plot to show top 5 GGG scores and bottom 5 GGG scores.
```{r}
GGG_plot = grid.arrange(p1, p2, ncol = 2)
print(GGG_plot)

ggsave("GGG_plot.png",
  plot = GGG_plot,
  dpi = 600,
)
```
Create a combined plot to show top 5 GII scores and bottom 5 GII scores.
```{r}
GII_plot = grid.arrange(p3, p4, ncol = 2)

ggsave("GII_plot.png",
  plot = GII_plot,
  dpi = 600,
)
```






